---
title: "Introduction to Data Analysis in R"
subtitle: "SSE Masters Course 7316"
output: 
  html_document:
    theme: flatly
    df_print: kable
    highlight: haddock
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
header-includes:
  - \usepackage{color} 
  - \usepackage{verbatim}
  - \usepackage{hyperref}
  - \definecolor{cornflower}{HTML}{1c86ee}
  - \hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    }   
---
```{r setup, include=FALSE}
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

<br>

## Course Description

This is an introduction to the R statistical programming language, focusing on
essential skills needed to perform data analysis from entry, to preparation, analysis,
and finally presentation.  During the mini-course, you will not only learn basic
R functionality, but also how to leverage the extensive community-driven package 
ecosystem, as well as how to write your own functions in R.

Course content is broken up into 7 seminars, each covering one content module 
except for the final review seminar. The length of each seminar may vary from 
module to module, but should generally be less than 3 hours.  The first hour or so
will be used to introduce new information, while the remainder of the 
time will be spent doing hands-on practice.  The content for each module will be 
posted the day before the seminar, so that you can familiarize yourself with the 
material ahead of time if you like (though this isn't required).

<br>

## Course Content

### **Module 1:**  Introduction to Base R Environment

This module introduces the R programming language and the RStudio software.  R 
programming topics will include coverage of basic operations
and data object types, especially vectors, matrices, and data frames.

- <font size="4"> **Lecture Notes**:</font>   \ \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/module1.html" role="button">Webpage</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/module1_slides.pdf" role="button">Slides</a>

- <font size="4"> **Seminar Exercises:**</font>   \ \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/setup_instructions.pdf" role="button">Setup Instructions</a> 
 <a class="btn btn-primary" href="https://www.datacamp.com/courses/6356" role="button">DataCamp Exercises</a>
 
```{r, include=FALSE}
#
```

<br>


### **Module 2:**  Data Preparation Using the Tidyverse

This module introduces a series of tools for data manipulation/preparation 
collectively known as the "Tidyverse."  Specifically, this module covers how to 
subset data, arrange it, transform it, and aggregate it.  Participants will also 
learn convenient tools to import and export data.

- <font size="4"> **Lecture Notes**:</font>   \ \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/module2.html" role="button">Webpage</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/module2.pdf" role="button">Slides</a>

```{r, include=FALSE}
```
  
- <font size="4"> **Seminar Exercises:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m2.pdf" role="button">Exercise</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/data/nlsy97.zip" role="button">nlsy97.zip</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m2-solutions.pdf" role="button">Solutions (PDF)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m2-solutions.R" role="button">Solutions (R Script)</a>

- <font size="4"> **Seminar Exercise 2018:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m2-2018.pdf" role="button">Exercise</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m2-solutions-2018.pdf" role="button">Solutions (PDF)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m2-solutions-2018.R" role="button">Solutions (R Script)</a>

<br>


### **Module 3:**  Programming, Joining Data, and More

This module introduces more advanced programming techniques to adapt R 
functionality to your own specific problems.  Contents include how to
perform loops, use conditional statements, and write basic functions.  In 
addition, this module will cover how to join data sets in R using 
Tidyverse functions, manipulating strings, and scrape tables from the web.

- <font size="4"> **Lecture Notes**:</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/module3.html" role="button">Webpage</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/module3.pdf" role="button">Slides</a> 

- <font size="4"> **Seminar Exercises:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m3.pdf" role="button">Exercise</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m3-solutions.pdf" role="button">Solutions (PDF)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m3-solutions.R" role="button">Solutions (R Script)</a>

- <font size="4"> **Seminar Exercises 2018:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m3-2018.pdf" role="button">Exercise</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m3-solutions-2018.pdf" role="button">Solutions (PDF)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m3-solutions-2018.R" role="button">Solutions (R Script)</a>

<br>


### **Module 4:**  Project Management and Dynamic Documents

This module provides a few major enhancements to the workflow process of data 
analysis in R.  Fist, Knitr and RMarkdown are introduced as a means to create 
dynamic reports from R using a variety of formats, such as HTML pages, PDF 
documents, and beamer presentations.  Then, RStudio Projects are introduced as 
means of organizing folders for empirical projects.  Finally, Git and GitHub are 
introduced to perform version control.

- <font size="4"> **Lecture Notes**:</font>   \ \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/module4.html" role="button">Webpage</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/module4.pdf" role="button">Slides</a> 

```{r, include=FALSE}


```

- <font size="4"> **Seminar Exercises:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m4.pdf" role="button">Exercise</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m4-pdf.pdf" role="button">Solutions (PDF)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m4-beamer.pdf" role="button">Solutions (Beamer)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m4-html.html" role="button">Solutions (html)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m4-solutions.Rmd" role="button">Solutions (Rmd)</a>

<br>


### **Module 5:** Regression Analysis and Data Visualization in R

In this module, standard linear regression in R is introduced, as 
well as coverage of common diagnostics and post-estimation procedures.  In addition,
further methods of regression analysis are covered, with special emaphasis on 
methods for panel and instrumental variables data.  Finally, the ggplot2 package 
is introduced as a means of creating compelling graphs in R.


- <font size="4"> **Lecture Notes**:</font>   \ \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/module5.html" role="button">Webpage</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/module5.pdf" role="button">Slides</a> 

- <font size="4"> **Seminar Exercises:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m5.pdf" role="button">Exercise</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m5b-extract.Rmd" role="button">Exercise -  Part B</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/data/nlsy97.rds" role="button">nlsy97.rds</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m5-parta-solutions.pdf" role="button">Solutions (PDF)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m5-parta-solutions.Rmd" role="button">Solutions (Rmd)</a>

- <font size="4"> **Seminar Exercises 2018:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m5-2018.pdf" role="button">Exercise</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/data/wdi_data.rds" role="button">wdi_data.rds</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m5-solutions-2018.pdf" role="button">Solutions (PDF)</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m5-solutions-2018.Rmd" role="button">Solutions (Rmd)</a>
```{r, include=FALSE}

```

 <br>
 
### **Module 6:** Introduction to Bayesian Methods in R 

This module introduces the basic intuition of Bayesian statistical methods and 
how to perform Bayesian analysis in R, primarily using the rstanarm package.

- <font size="4"> **Lecture Notes**:</font>   \ \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/module6.html" role="button">Webpage</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/module6.pdf" role="button">Slides</a>
```{r, include=FALSE}

```

- <font size="4"> **Seminar Exercises:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/exercise-m6.pdf" role="button">Exercise</a>

```{r, include=FALSE}

```

 <br> 

### **Module 7:**  Review Seminar for Capstone Project

In this module, an extended empirical exercise is utilized to review the skills 
developed over the preceding seminars.  The review will function as preparation 
for the capstone project, in which students individually replicate results from 
a recent economics paper.

- <font size="4"> **Seminar Exercises:**</font>   \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/review_exercise.pdf" role="button">Exercise</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/datanames.csv" role="button">datanames.csv</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/review_exercise-solutions.pdf" role="button">Solutions (PDF)</a> 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/review_exercise-solutions.Rmd" role="button">Solutions (Rmd)</a> 
```{r, include=FALSE}




#
```

<br>
<br>

### **Capstone Project:**  Review Seminar for Capstone Project
For the capstone project, you will be replicating results from 
"Intergenerational Mobility and Preferences for Redistribution" (AER 2018) by 
Alberto Alesina, Stefanie Stantcheva, Edoardo Teso.  The capstone project is 
due on March 18^th^.

-  <font size="4"> **Files:**</font> \ \ \ \ 
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/capstone.pdf" role="button">Instructions</a>
<a class="btn btn-primary" href="https://andrewproctor.github.io/rcourse/assets/actual_probabilities" role="button">actual_probabilities.csv</a>

## Contact:

   **Course Teacher:**  Andrew Proctor
  
   **Office:**  A 711 (Arrange by email or stop in if I'm there.)
  
   **Email:**  andrew.proctor@phdstudent.hhs.se

