---
title: "Introduction to R"
subtitle: "Module 2: Data Preparation using the Tidyverse"
institute: "andrew.proctor@phdstudent.hhs.se"
date: "February 5, 2018"
output: 
  html_document:
      theme: flatly
      highlight: haddock
      toc: true
      toc_float: true

---

```{r setup, include=FALSE}
library(tidyverse)

```

<a class="btn btn-primary" href="https://andrewproctor.github.io/rminicourse/assets/module2-dataprep.pdf" role="button">Slides</a>

<br>

# Intro

## Goals for Today

1. Learn how to use packages in R
2. Learn how to import and export data.
3. Learn how to perform common data prep functions from the tidyverse 
collection of packages.
4. Learn how to clean and "tidy" data.

# Packages in R

## Role of Packages in R

- Packages in R are similar to user-written commands (think *ssc install*) in Stata.

- But *most* things you do in Stata probably use core Stata commands.

- In R, most of your analysis will probably be done using packages.

\includegraphics[scale=0.4]{cran.png}

## Installing and using a package

- To install a package, use the function (preferably in the console) [install.packages()](https://www.rdocumentation.org/packages/utils/versions/3.4.3/topics/install.packages)



- To begin with, let's install 2 packages:
    + [tidyverse](https://www.rdocumentation.org/packages/tidyverse):  the
    umbrella package for common data preparation and visualization in R.
    + [rio](https://www.rdocumentation.org/packages/rio):  a package for easy
    data import, export (saving), and conversion.

```{r , eval=FALSE}
install.packages("tidyverse")    # Install tidyverse
install.packages("rio")    # Install rio
```    
    
## Loading a package during analysis

Unlike Stata, in R you need to declare what packages you will be using at the
beginning of each R document.

To do this, use the [library()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/library) 
function.
    _ [require()](https://www.rdocumentation.org/packages/tidyverse) also 
    works, but it's use is discouraged for this purpose.

```{r, warning= FALSE, message=FALSE}
library("tidyverse")    # Install tidyverse
library("rio")    # Install rio
```  

# Data Prep Preliminaries

## Import and export using rio

Previously, importing and exporting data was a mess, with a lot of different 
functions for different file formats:  
  
- Stata DTA files alone required two functions:   read.dta (for Stata 6-12 
DTA files),  read.dta13 (for Stata 13 and later files), etc.

The [rio](https://www.rdocumentation.org/packages/rio) package simplifies 
this by reducing all of this to just one function, 
[import()](https://www.rdocumentation.org/packages/rio/versions/0.5.5/topics/import)

- Automatically determines the file format of the file and uses the appropriate
function from other packages to load in a file.

```{r, warning= FALSE}
PISA_2015 <- import("PISA2015.sas7bdat")
PISA_2015[1:5,1:6]

export(PISA_2015, "PISA_2015.rds")
``` 


## Tibbles: an update to the data frame

Last class, we covered data frames---the most basic data object class for data 
sets with a mix of data class.

Today, we introduce one final data object: the **tibble**!

The tibble can be thought of as an update to the data frame---and it's the first 
part of the *tidyverse* package that we'll look at.

### Tibble vs data frames

There are three main benefits to the tibble:

1.  Displaying data frames:
    + If you display a data frame, it will print as much as much output as allowed 
    by the "max.print" option in the R environment.  With large data sets, that's 
    far too much. Tibbles by default print the first 10 rows and as many columns 
    as will fit in the window.

2. Partial matching in data frames:  
    + When using the <span style="color:blue">**$**</span> method to reference columns of a
    data frame, partial names will be matched if the reference isn't exact.  This
    might sound good, but the only real reason for there to be a partial match
    is a typo, in which case the match might be wrong.
    
3. Tibbles are required for some functions.

### Creating or converting to tibbles
The syntax for creating tibbles exactly parallels the syntax for data frames:

- [tibble()](https://www.rdocumentation.org/packages/tibble) creates a 
tibble from underlying data or vectors.

- [as.tibble()](https://www.rdocumentation.org/packages/tibble/versions/1.4.2/topics/as_tibble)
coerces an existing data object into a tibble.

```{r, warning= FALSE}
PISA_2015 <- as.tibble(PISA_2015); PISA_2015[1:5,1:5]
``` 

## Glimpse

Another tidyverse function that's very useful is 
[glimpse()](https://www.rdocumentation.org/packages/tibble/versions/1.4.2/topics/glimpse) \href{}{}, a function very similar to 
[str()](https://www.rdocumentation.org/packages/utils/versions/3.4.3/topics/str).

- Both functions display information about the structure of a data object.

- [str()](https://www.rdocumentation.org/packages/utils/versions/3.4.3/topics/str) 
provides more information, such as column (variable) attributes embedded 
from external data formats, but consequently is much less readable for complex 
data objects.

- [glimpse()](https://www.rdocumentation.org/packages/tibble/versions/1.4.2/topics/glimpse) 
provides only column names, classes, and some data values (much more
readable)

- I will often use [str()](https://www.rdocumentation.org/packages/utils/versions/3.4.3/topics/str) 
when I want more detailed information about data structure, but use [glimpse()](https://www.rdocumentation.org/packages/tibble/versions/1.4.2/topics/glimpse) 
for quicker glances at the data.


## Pipes

Another major convenience enhancement from the tidyverse is ***pipes***, denoted
<span style="color:blue">**%>%**</span>,

- Pipes allow you to combine multiple steps into a single piece of code.

- Specifically, after performing a function in one step, a pipe takes the data 
generated from the first step and uses it as the data input to a second step.

### Pipes Example

```{r}

barro.lee.data <- import("BL2013_MF1599_v2.1.dta") %>% 
  as.tibble() %>% glimpse(width = 50)
```

# Data Preparation

## Tidyverse and the verbs of data manipulation

A motivating principle behind the creation of the tidyverse was the language of
programming should really behave like a language.

Data manipulation in the tidyverse is oriented around a few key "verbs" that 
perform common types of data manipulation.

1. [filter()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/filter) 
subsets the rows of a data frame based on their values.
2. [select()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/select) 
keeps variables (columns) based on their names.
3. [mutate()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/mutate) 
adds new variables that are functions of existing variables.
4. [summarize()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/mutate) 
creates a number of summary statistics out of many values.
5. [arrange()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/arrange) 
changes the ordering of the rows.

**Note**:  the first argument for each these functions is the data \ \ \ 
object (so pipe!).

## Filtering data

Filtering keeps observations (rows) based on conditions.

- Just like using use subset conditions in the row arguments of a bracketed subset

```{r, include=FALSE}
library(Ecdat)
data(Males)
set.seed(28133)
wages <- Wages1 %>% as.tibble() %>% sample_n(3) %>% as.data.frame() %>% 
  select(wage,school, sex, exper) %>% rename(schooling = school) %>% 
  mutate(wage=exp(wage))

# Source: Verbeek, Marno (2004) A Guide to Modern Econometrics, John Wiley and Sons.
```

```{r}
# Using brackets
wages[(wages$schooling > 10) & (wages$exper > 10),] 
# Using filter
wages  %>% filter(schooling > 10,exper > 10)  


```

Notice a couple of things about the output:

1.  It doesn't look like we told [filter()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/filter)
what data set we would be filtering.
+ That's because the data set has already been supplied by the pipe.  We could
have also written the filter as:
```{r}
filter(wages, schooling > 10,exper > 10)  
```

2.  We didn't need to use the logical <span style="color:blue">**&**</span>. Though 
multiple conditions can still be written in this way with filter(), the default
is just to separate them with a comma.

## Selecting data

Just like filter is in many ways a more convenient form of writing out bracketed
row subset conditions, the verb 
[select()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/select)
is largely a more convenient method for writing column arguments.

```{r, include=FALSE}
wages_row1 <-wages[1,]

```

```{r}
# Using brackets
wages_row1[,c("wage","schooling","exper")]
# Using select
wages_row1  %>% select(wage,schooling,exper)  
```

### An example of dropping a column

One option we have not covered so far in creating subsets is dropping rows or columns.

R has a specific notation for this, easily used with 
[select()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/select):
```{r}
wages_row1 # What wages_row1 looks like:
wages_row1  %>% select(-exper) #drop exper
```

Dropping columns (or rows) using the <span style="color:blue">**-**</span> notation also 
works with brackets, but only when using the number location of the row or 
column to be dropped.

```{r}
wages_row1[,-4] # works
# wages_row1[,-"exper"] does not work
```

Because of the ability to use name arguments, dropping with 
[select()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/select)
is generally easier.

## "Mutating" data

Creating new variables that are functions of existing variables in a data set 
can be done with [mutate()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/mutate).

[mutate()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/mutate) 
takes as its first argument the data set to be used and the equation for the new
variable:

```{r}
wages <- wages %>% 
  mutate(expsq = exper^2) # Create expersq
wages # Display wages
```

## Summarizing data

Summary statistics can also be easily created using the tidyverse function 
[summarize()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/mutate) 

The [summarize](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/summarise) 
functions uses summary statistic functions in R to create a new summary tibble, with syntax
largely identical to mutate().

Let's try summarizing with the [mean()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/mean) 
summary statistic.

```{r}
wages %>% 
  summarize(avg_wage = mean(wage))
```


### Summary Statistics functions in R

There are a number of summary statistics available in R, which can be used 
either with the [summarize()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/summarise)
command or outside of it:

**Measures of central tendency and spread:**

- [mean()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/mean),
[median()](https://www.rdocumentation.org/packages/stats/versions/3.4.3/topics/median),
[sd()](https://www.rdocumentation.org/packages/stats/versions/3.4.3/topics/sd),
[var()](https://www.rdocumentation.org/packages/stats/versions/3.4.3/topics/cor),
[quantile()](https://www.rdocumentation.org/packages/stats/versions/3.4.3/topics/quantile),
[IQR()](https://www.rdocumentation.org/packages/stats/versions/3.4.3/topics/IQR) 

**Position:**

- [first()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/nth),
[last()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/nth),
[nth()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/nth),

**Count:** 

- [n()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/n),
[n_distinct()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/n_distinct),

### Multiple summary variables

Let's look at an example of using multiple summary variables with a larger 
50-observation sample for the `wages` data set.

```{r, include=FALSE}
library(Ecdat)
data(Males)
set.seed(1)
wages <- Wages1 %>% as.tibble() %>% select(wage,school, sex, exper) %>% 
  rename(schooling = school) %>% filter(wage > 0) %>% filter(wage < 13) %>% 
  sample_n(30) %>% mutate(wage=exp(wage))

# Source: Verbeek, Marno (2004) A Guide to Modern Econometrics, John Wiley and Sons.
```

```{r}
wages %>% 
  summarize(avg.wage = mean(wage), sd.wage = sd(wage), 
        avg.exper = mean(exper), sd.exper = sd(exper))
```


## Grouping data

Creating summary statistics by group is another routine task.  This is 
accommodated in the tidyverse using the [group_by()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/group_by).

- The arguments of group_by(), in addition to the data set, are simply the grouping
variables separated by commas.

```{r}
wages %>% group_by(sex) %>%
  summarize(avg.wage = mean(wage), sd.wage = sd(wage))
```

## Arranging (sorting) data

If you want to sort your data by the values of a particular variable, you can 
easily do so as well with the 
[arrange()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/arrange)  
function.

```{r}
wages[1:3,] %>% arrange(exper)
```

**Not:** [arrange()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/arrange) 
sorts values in ascending order by default.  If you want to sort in descending 
order, wrap the variable name inside <span style="color:blue">**desc()**</span> in the function.

## Sampling from data

Creating a sample from a data set in R is made easy by two main function in R: 
[sample_n](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/sample) and [sample_frac](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/sample).

**Syntax:**

- sample_n(data, size, replace = FALSE/TRUE)

- sample_frac(data, size = 1, replace = FALSE/TRUE)

## A data prep example with fuel economy data

Let's use tidyverse data manipulation verbs to work through a practical data prep 
problem from start to finish.

For the problem, Let's use fuel economy data again, but with half of the data set. 
The data comes from the `vehicles` data set in the 
[fueleconomy](https://cran.r-project.org/web/packages/fueleconomy/index.html) 
package.


```{r}
# install.packages("fueleconomy") # Run only once
library(fueleconomy)
```

Now let's look at how fuel efficiency has changed over time in the data set. 
Specifically, let's create descriptive statistics of fuel efficiency by year for 
"normal" passenger vehicles (4-8 cylinders).

### What's in the data set?

```{r}
glimpse(vehicles[2:12], width=50)
```

### Create summary tibble

```{r}

annual.mpg <- vehicles %>% sample_frac(0.5) %>% 
  filter(cyl %in% 4:8) %>% group_by(year) %>%  
  summarize(hwy.avg = mean(hwy), hwy.sd = sd(hwy),
            city.avg = mean(cty), city.sd = sd(cty)) %>% 
  arrange(desc(city.avg))
```

**Note:** Here I used <span style="color:blue">**%in%**</span>, which works like 
<span style="color:blue">**inrange**</span> in Stata.  You could alternately write two 
inequalities to achieve the same thing.

```{r}
# Print annual.mpg
annual.mpg
```

## Summarizing a data set with the summary() function

Although the tidyverse [summarize()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/summarise) 
function is more powerful, often you just a want a quick look at summary statistics 
for the whole data set.

- You can easily do this with the base R
[summary()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/summary) 
function, which produces summaries not just for data sets, but also for other R output like the 
results of a regression.

```{r}
summary(wages)

```


# Cleaning data


## Common data cleaning tasks

There are a few data cleaning tasks that are pervasive in empirical work:

1. Ensure columns have useful names
2. Recoding variable values
3. Addressing missing values

## Renaming columns

Renaming columns is easily accommodated with the tidyverse 
 [rename()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/select) \href{}{}
command.

- Arguments are the data set and *NewVarName* = *OldVarName*.

To see [rename()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/select) 
in action, let's go back to the `barro.lee.data` educational data set we imported earlier:

### Renaming columns example

Let's look at columns 1 and 7 through 9:
```{r}
glimpse(barro.lee.data[,c(1,7:9)], width = 50)
```

See how these variable names are uninformative?  Let's use str() now.

```{r}
str(barro.lee.data[,c(1,7:9)])
```

**Although** the column names here are uninformative, the imported SAS file
included variable labels saved as attributes, which can be seen using the more
verbose [str()](https://www.rdocumentation.org/packages/utils/versions/3.4.3/topics/str) 
function.

```{r}
barro.lee.data <- barro.lee.data %>% 
  rename(countrycode = BLcode, 
    perc.noschool = lu, 
    perc.primary = lp,
    perc.primary.complete = lpc)

```

Now let's look at the variable names again:

```{r}
glimpse(barro.lee.data[,c(1,7:9)], width = 50)
```

## Recoding variables

Along with renaming variables, recoding variables is another integral part of 
data wrangling.  

```{r}
wages[1:4,"sex"] # Look at sex column

wages$sex <- wages$sex %>% recode("male"=0, 
                            "female"=1) # recode
wages[1:4,"sex"] # Look at sex column

```

## Missing Values

Another problem characteristic of observational data is missing data.  In R,
the way to represent missing data is with the value <span style="color:blue">**NA**</span>.

- You can recode missing value that *should be* NA but are code using a different 
schema either by using brackets, or the tidyverse 
[na_if()](http://dplyr.tidyverse.org/reference/na_if.html) function.

```{r, eval=FALSE}
## Replace 99-denoted missing data with NA
# bracket method
wages[wages$schooling==99,] <- NA  
# tidyverse method
wages$schooling <- wages$schooling %>% na_if(99) 
```

You can check for (correctly-coded) missing-values using the 
[is.na()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/NA) 
function.

```{r, include=FALSE}
## Replace 99-denoted missing data with NA
wages[c(1,8,21),"wage"] <- NA

```

```{r}
## Missing 
wages[is.na(wages$wage),]

```

**Note:** R does not naturally support multiple types of missingness like other 
languages, although it's possible to use the [sjmisc](https://cran.r-project.org/web/packages/sjmisc/index.html) 
package to do this.

# Tidy data

## Principles of tidy data

Rules for tidy data (from *R for Data Science*):

1. Each variable must have its own column.
2. Each observation must have its own row.
3. Each value must have its own cell.

## Tidy data tools in the tidyverse

There two main tidyverse verbs for making data tidy are:

**[gather()](https://www.rdocumentation.org/packages/tidyr/versions/0.8.0/topics/gather)**:  reduces variable values are spread over multiples columns into a single column.

**[spread()](https://www.rdocumentation.org/packages/tidyr/versions/0.8.0/topics/spread)**:  when multiple variables values are stored in the same columns, moves each variable into it's own column.


## Gathering data

If values for a single variable are spread across multiple columns (e.g. income 
for different years), [gather](https://www.rdocumentation.org/packages/tidyr/versions/0.8.0/topics/gather) 
moves this into single "values" column with a "key" 
column to identify what the different columns differentiated.

**Syntax:** gather(*data*, *key*, *value*, *columnstocombine*)

### Gather example
```{r, include = FALSE}

person <- c("Elsa","Mickey","Ariel","Gaston","Jasmine","Peter","Alice")
y1999 <-c(10,20,17,19,32,22,11)
y2000 <-c(15,28,21,19,35,29,15)
earnings.panel <-tibble(person,y1999,y2000)

```

```{r}
earnings.panel

earnings.panel <- earnings.panel %>% 
  gather(key="year", value="wage",y1999:y2000)
earnings.panel
```

## Spreading data

Spread tackles the other major problem - that often times (particularly in 
longitudinal data) many variables are condensed into just a "key" (or indicator) 
column and a value column.

```{r, include = FALSE}
wages2 <- earnings.panel$person %>% cbind(wages[1:7,c(1,2,4)])  %>%
  gather(key="indicator",value="values", 2:4)

colnames(wages2)[1] <- "person"

```

```{r}
wages2

```

```{r, eval=FALSE}
wages2 %>% spread("indicator", "values")

```

```{r, include = FALSE}
wages2 <- earnings.panel$person %>% cbind(wages[1:7,c(1,2,4)])
colnames(wages2)[1] <- "person"

```

```{r, echo = FALSE}
wages2
```
