---
title: "Module 7 (Bayesian) Exercise"
output: 
  pdf_document:
    df_print: kable
    highlight: zenburn
urlcolor: blue
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstanarm)
library(rio)
library(tidyverse)
library(rstan)
library(bayesplot)
```

## Preliminaries

- Install RTools  from 
[here](https://github.com/stan-dev/rstan/wiki/Installing-RStan-on-Windows):

- Install and initialize the packages rstan, rstanarm, and shinystan.  

    + Be sure to initialize any other packages you would normally use.
    
- In a new or existing RStudio Project, create an RMarkdown document.

## Analysis of Well Switching

- Check out the descriptions of the datasets included with the rstanarm package [here](https://www.rdocumentation.org/packages/rstanarm/versions/2.17.2/topics/rstanarm-datasets):


- Using the **wells** dataset, perform a Bayesian linear regression that uses arsenic level, 
distance from well, and years of education to predict well-switching.
    
```{r, message=FALSE}
wellsLM <- stan_glm(switch ~ arsenic + 
    dist + educ, 
    data = wells, family = gaussian)

```

- View the summary of the regression output.

```{r}
summary(wellsLM)
```

- Create a histogram of the posterior distribution of the *dist* variable

```{r}
wells_posterior <- wellsLM %>% as.tibble()

ggplot(wells_posterior, aes(x=dist)) + geom_histogram()
```

- Use the leave-one-out method of measuring out-of-sample predictivenes to compute 
the expected log predictive density.

```{r}
loo(wellsLM)


```

## Analysis of doctors visits

- If necessary, install the package "Ecdat" and initialize it.

- Using the dataset **Doctor**, perform a Bayesian Poisson regression that 
estimates the number of doctor visits as a function of health care access and health 
status.

```{r, message=FALSE}
library(Ecdat)
doctordata <- Doctor

doctorvisits_poissonreg <- stan_glm(doctor ~ access + 
    health, 
    data = doctordata, family = poisson)



```

- Once again view the summary of the output but this time view a graphical plot 
of the credible intervals for *access*.

```{r}

doctorvisits_posterior <- doctorvisits_poissonreg %>% as.tibble()

ggplot(doctorvisits_posterior, aes(x=access)) + geom_histogram()

```

- Use the Watanabe-Aikake information criterion to view the predictiveness of the 
Poisson model.

```{r}

waic(doctorvisits_poissonreg)

```


