---
title: "Exercise 3"
output: 
  beamer_presentation:
      theme: "CambridgeUS"
      colortheme: "orchid"
      highlight: espresso
graphics: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
library(rvest)
library(httr)
library(magrittr)
library(tidyverse)
library(rio)
```

For this exercise, you will not be working with a single dataset, but will 
instead practice new skills using both your NLSY data from last seminar, as well 
as online data concerning movies.

## 1.  Revisit your NLSY97 dataset from last week
    
  a. Create an indicator for sex using a vectorized conditional statement.
  b. Recode the schooltype variable into text values, corresponding to:
      
    - "Public" if the value is 1
    - "Private, religious" if the value is 2
    - "Private, non-religious" if the value is 3
    - "Other" if the value is 4.

```{r}
    nlsy97 <-import("nlsy97.rds")
```

<br>

# 2. Load the IMDB Top 250 Movies

## 2a: Scrape the data from the "Top 250 Movies as rated by IMDb users

From [https://www.imdb.com/chart/top](https://www.imdb.com/chart/top):
    
```{r}
# IMDB Top 250 Movies
top250_basic <- 
  read_html("https://www.imdb.com/chart/top/") %>%
  html_table() %>% as.data.frame()
```
    
## 2b. Notice that IMDB scrapes the data in Swedish by default.  

To get the data in English, use **html_session()** in place of **read_html()**, 
adding the option:
      
```{r, eval=FALSE, include=TRUE}
    add_headers("Accept-Language"="en-US, en;q=0.5")
```
  
  - You may need to load the **httr** package to use **add_headers()**.
      
    ```{r}
top250_eng.pre <- 
      html_session("https://www.imdb.com/chart/top/", 
      add_headers("Accept-Language"="en-US, en;q=0.5")) %>% 
      html_table %>% as.data.frame()
    ```
    
## Steps 2c-d

**2c:  Keep only the columns "Rank...Title" and "IMDb.Rating", suitably renaming them.**
    
```{r}
top250_eng <- top250_eng.pre %>% 
  select("Rank...Title", "IMDb.Rating") %>%
  rename(Title = Rank...Title,
  Rating = IMDb.Rating) # Keep only Title and Year Columns
```
    
**2d. Create a ranking variable by extracting the values that appear before the 
dot in the title column.**

```{r}
top250_eng$Ranking <- top250_eng$Title %>% 
  str_extract("[0-9]+(?=(.\n))")
```

## 2e-2g

**2e. Create a year variable, by extracting the numbers inside a parenthesis 
from the title column.**

```{r}
top250_eng$Year <- str_extract(top250_eng$Title, 
                      "(?<=[:punct:])[:digit:]+") 
```

**2f. Redefine the title variable by extracting the string information that appear after the dot in the title column.**
    
```{r}
 top250_eng$Title <- top250_eng$Title %>% 
  str_extract("(?<=(.\n)).+")
```

**2g. Trim the white space on both sides of the title.**

```{r}
top250_eng$Title %<>% str_trim(side = "both")
```

# Important Tables and Figures

## View Top 250 dataframe

```{r, results = "hide"}
head(top250_eng)
```

Invisible table!

## Elite R Programmer

Keep working and one day soon, this can be you!

\includegraphics{rcat.jpg}

