---
title: "Module 5 Exercise - Part B"
author: "Your Name Here"
output: 
  pdf_document:
    df_print: kable
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following is an exercise where the first two (data preparation) steps are 
already done for you.  Complete steps 3 and 4 to visualize the data and 
perform regression analysis.

## 1. Create an extract of the nls97 dataset.

- Initialize the packages used in this analysis.
    ```{r, message = FALSE, warning = FALSE}
    library(rio)
    library(tidyverse)
    library(plm)
    library(ggplot2)
    library(ggthemes)
    library(broom)
    library(lmtest)
    library(stargazer)
    ```

- Load the final nlsy97 dataset extract (taken from the end of Module 2).
    ```{r}
    nlsy97 <- import("nlsy97.rds")
    ```

- Create the following new variables:

    + *logparentincome*, equal to the log of parent income.
    + A variable for the highest degree completed by the mother, equal to:
        - College if *motheredyrs* is greater than or equal to 14.
        - High school if *motheredyrs* is between 12 and 13.
        - Less than high school if *motheredyrs* is less than 12.
    ```{r}
    nlsy97 <- nlsy97 %>% mutate(logparentincome = log(parentincome))
    
    nlsy97 <- nlsy97 %>% mutate(mother_degree = case_when(
      motheredyrs >= 14 ~ "college",
      motheredyrs %in% 12:13 ~ "high school",
      motheredyrs %in% 0:11 ~ "less than high school"
    ))
    ```

- Turn the mother's degree variable into an ordered factor.
    ```{r}
    nlsy97 <- nlsy97 %>% 
      mutate(mother_degree = factor(mother_degree, 
             levels = c("less than high school",
                        "high school", 
                        "college"),
             ordered=TRUE))
    ```

## 2. Create a new dataset, **nlsy97_sample**, modified from **nlsy97**, which:

- Drops missing values in the following variables:
  
    + *parentincome*
    + *motheredyrs*
    + *gpa*
    + *highestgrade*
    
    ```{r}
    nlsy97_sample <- nlsy97 %>% drop_na(parentincome,motheredyrs,gpa,highestgrade)
    # Another approach would be:
    # nlsy97 <- nlsy97 %>% filter(!is.na(parentincome),
    #                             !is.na(motheredyrs),
    #                             !is.na(gpa),
    #                             !is.na(highestgrade))
    ```

- Subsets the dataset for only observations where the student is 18 years old,
with a GPA greater than or equal to 1.3 (D+ average), with parent income 
between $5,000 and $100,000.

    ```{r}
    nlsy97_sample <- nlsy97_sample %>% filter(age == 18,
                                gpa > 1.3,
                                parentincome > 5000, 
                                parentincome < 100000)
    ```

## 3. Create a scatterplot

Create a scatterplot with:

- The dataset is 75 randomly sampled observations from **nlsy_sample** 
- Log Parental income is on the x-axis, shown from values of 8 to 12.
- GPA is on the y-axis, shown from values 1 to 4.
- The color of points based on the mother's years of education.
- The color scale set using the following:
    ```{r, eval=FALSE}
    scale_colour_brewer(palette = "Set1")
    ```
- Add a title and axis labels.

```{r, message=FALSE, warning=FALSE}
ggplot(sample_n(nlsy97_sample,75), 
       aes(x = logparentincome, y = gpa, col= mother_degree)) + 
  geom_point() + scale_colour_brewer(palette = "Set1") + 
  ggtitle("Intergenerational Education and Income") +
  xlab("Parental Income (logged)") + ylab("GPA") + 
  ylim(1,4) + xlim(8,12) + theme_hc()
```

## 4. Perform a regression analysis of the effect of parent income on GPA
- Remove observations with Parental Income less than $5,000 and GPA less than or equal to 1.
```{r}
nlsy97 <- nlsy97 %>% filter(parentincome>5000, gpa > 1)
```

- First run an OLS regression, with:
    + The log of GPA as the dependent variable
    + School type and log of parent income as the independent vaariables

```{r}
gpa_and_parentincome <- lm(log(gpa) ~ log(parentincome) + schooltype, data = nlsy97)
```
- Display the results with tidy
```{r, results = "asis"}
tidy(gpa_and_parentincome)
```

- Test for autocorrelation and report the statistical decision.
```{r}
bgtest(gpa_and_parentincome)
```

- Re-run the regression as fixed effects regression, including both unit and time 
fixed effects.
```{r}
gpa_and_parentincome_fe <- plm(log(gpa) ~ log(parentincome) + schooltype, data = nlsy97, 
                        index = c("personid","year"), 
                        model = "within", effect="twoway")
```

- Perform autocorrelation-robust inference using the fixed effects regression

    +  Test the coefficients of the model using the Stata-style $HC_1$ 
    estimation of Newey-West heteroskedasticity and autocorrelated (HAC) 
    robust standard errors.
    
    + Use the function 
    [vcovNW()](https://www.rdocumentation.org/packages/plm/versions/1.6-5/topics/vcovNW) 
    for specifying the variance method inside of 
    [coeftest()](https://www.rdocumentation.org/packages/lmtest/versions/0.9-35/topics/coeftest).
    
    + View the results with **tidy()**

```{r}
tidy(coeftest(gpa_and_parentincome_fe,vcov = 
           vcovNW(gpa_and_parentincome_fe,type="HC1")))

```

