---
title: "Module 6: Practice with Bayesian Regression Analysis"
output: 
  pdf_document:
    df_print: kable
    highlight: zenburn
urlcolor: blue
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
library(rstanarm)
library(rio)
library(tidyverse)
library(rstan)
library(bayesplot)
```

\newcommand{\andrewcomment}[1]{{\leavevmode\color{blue}#1}}

\andrewcomment{\textbf{Note: } Completion of this exercise is voluntary - you 
are not required to submit this assignment. 
}

## Preliminaries

- If you are on Windows, you may first need to install RTools from 
[here](https://cran.r-project.org/bin/windows/Rtools/):

- Install and initialize the packages rstan, rstanarm, and shinystan.  

    + Be sure to initialize any other packages you would normally use.
    
- In a new folder of your project for the course, create a new pdf-type Rmarkdown
document.

## Analysis of Well Switching

- Check out the descriptions of the datasets included with the rstanarm package [here](https://www.rdocumentation.org/packages/rstanarm/versions/2.17.2/topics/rstanarm-datasets):


- Using the **wells** dataset, perform a Bayesian linear regression that uses arsenic level, 
distance from well, and years of education to predict well-switching.
    
    + Try repeating the regression using both normal(0) and laplace(0) priors, 
    viewing the summary of the regression output for each.
    
```{r, results = "hide"}
wellsLM <- stan_glm(switch ~ arsenic + 
    dist + educ, data = wells, 
    family = binomial(link=logit),
    prior = normal(0, autoscale = TRUE))
```

```{r}
summary(wellsLM)
```

```{r, results="hide"}
wells_laplacep <- stan_glm(switch ~ arsenic + 
    dist + educ, data = wells, 
    family = binomial(link=logit),
    prior = laplace(0))
```

```{r}
summary(wells_laplacep)
```

- Create a histogram of the posterior distribution for the *dist* variable from 
the regression using normal(0) priors.

```{r}
wells_posterior <- wellsLM %>% as.tibble()

ggplot(wells_posterior, aes(x=dist)) + geom_histogram()
```

- Use the leave-one-out method of measuring out-of-sample predictivenes to compute 
the expected log predictive density.

```{r}
loo(wellsLM)
```

## Analysis of doctors visits

- If necessary, install the package "Ecdat" and initialize it.

- Using the dataset **Doctor**, perform a Bayesian Poisson regression that 
estimates the number of doctor visits as a function of health care access and 
health status.

    + Perform the regression for a sample of 100 observations, for each of (1) 
    the default priors and (2) uninformative priors.

```{r, results = "hide"}
library(Ecdat)
doctordata <- Doctor
doctordata <- doctordata %>% sample_n(100)

doctorvisits_poissonreg <- stan_glm(doctor ~ access + 
    health, data = doctordata, family = poisson, prior = normal(1))
```

```{r}
summary(doctorvisits_poissonreg)
```


```{r, results = "hide"}
doctorvisits_poissonreg_uninfprior <- stan_glm(doctor ~ access + 
    health, 
    data = doctordata, family = poisson, prior = NULL)
```

```{r}
summary(doctorvisits_poissonreg_uninfprior)
prior_summary(doctorvisits_poissonreg_uninfprior)
```
- Once again view the summary of the output but this time view a graphical plot 
of the credible intervals for *access*.

```{r}

doctorvisits_posterior <- doctorvisits_poissonreg %>% as.tibble()

ggplot(doctorvisits_posterior, aes(x=access)) + geom_histogram()

```

- Use the Watanabe-Aikake information criterion to view the predictiveness of the 
Poisson model.

```{r}
waic(doctorvisits_poissonreg)
```


