---
title: "Module 5 Exercise"
output: 
  pdf_document:
    df_print: kable
    highlight: tango
urlcolor: blue

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
# Initialize libraries
library(tidyverse)
library(rio)
library(plm)

```

## 1. Preliminaries

### a. Create a Project and RMarkdown document 

Either in a preexisting RStudio Project folder, or in a new Project, create a 
RMarkdown document for the exercise.

### b. Initialize the Ecdat library

For the first part of the exercise, you will be working with the [Ecdat](https://www.rdocumentation.org/packages/Ecdat/) package, 
which has many economics datasets.

- Install and initalize the package.

```{r}
library(Ecdat)
```

## 2. Extramarital Affairs Exercise

### b. Create an "affairs" tibble from the Fair data

For the first part of the exercise, you will be working the extramaritals data from 
Fair (Econometrica 1977).  

- Create an "affairs" data frame from the Fair data in Ecdat as follows:

```{r}
affairs <- Fair 
```

- Turn the data set into a tibble
- Select the following variables: `nbaffairs, ym, child, rate, age`.
- Rename the variables as follows:
    
    + Change "nbaffairs" to "affairs"
    + Change "ym" to "yrsmarr"
    + Change "rate" to "mrating"

```{r}
library(Ecdat)
affairs <- Fair %>% as.tibble() %>% select(nbaffairs, ym, child, rate, age) %>% 
  rename(affairs = nbaffairs, yrsmarr= ym, mrating = rate) 
```

### c. Convert mrating to a factor

```{r}
affairs$mrating <- ordered(affairs$mrating)
```


### d. Perform regression 

- Create a regression object that regresses affairs on age, child, yrsmarr and 
$yrsmarr^2$.

```{r}

affairs_reg <- lm(affairs ~ age + child + yrsmarr + I(yrsmarr^2), data = affairs)

```

- Then view the summary output of the regression.

```{r}
summary(affairs_reg)
```


### f. Perform model diagnostics

- **Test for heteroskestacity.** 
- **Test for missing polynomial terms.** 

Write the statistical decision from each test in your RMarkdown report.

```{r}
library(lmtest)
bptest(affairs_reg)
resettest(affairs_reg)
```

### g. Create a bar plot of average number of affairs by marital rating

- Create a summary data frame of number of affairs by marital rating
- Create a bar plot of the results, adding a graph title, axis labels, and 
style it like a Stata graph using **theme_stata()** from the [ggthemes](https://cran.r-project.org/web/packages/ggthemes/vignettes/ggthemes.html) 
package.

```{r}
library(ggthemes)

ggplot(affairs,
aes(x=mrating, y=affairs)) +
geom_bar( stat = "summary", fun.y = "mean") +
ggtitle("Affairs by Marital Rating") + 
xlab("Marital Rating [1-5]") + ylab("# of Affairs") + 
theme_stata()

```

## 3. Inequality in the World Development Indicators Example

For the next part of the exercise, let's revisit the inequality indicators from 
the World Development Indicators dataset.  

- Import "wdi_data", which is just the data from Exercise 2, summarized by 
country and year from 2000.

```{r}
wdi_data <- import("C:/Users/AN.4271/OneDrive - Handelshögskolan i Stockholm/Teaching/R Course/Modules/Regression Modelling/Exercise/wdi_data.rds")
```


### a. Create scatter plots 

- Create twos scatterplots: each with secondary school completion as the x-axis.
    + For the first scatterplot, choose female-to-male employment ratio as the 
    y-variable. 
    + For the second scatterplot, make unemployment among those with basic education 
    the y-variables:

- Color the points based on the value of GDP.

- Add titles and axis labels.

```{r}
ggplot(wdi_data, aes(x=secondary.complet.avg,y =unemp.basic.educ.)) + 
  geom_point(aes(color=GDP.avg))

ggplot(wdi_data, aes(x=secondary.complet.avg,y =femaletomale.avg)) + 
  geom_point(aes(color=GDP.avg))

```


### b. Perform regression

Perform a pooled OLS regression of GDP average on secondary school completion and 
view summary information from the regression.

```{r}
WDI_regs <- lm(GDP.avg ~ secondary.complet.avg, data= wdi_data)
summary(WDI_regs)

```

### c. Test for autocorrelation

```{r}
bgtest(WDI_regs)
```

### d. Fixed effects regression

Now instead perform the regression of GDP on secondary school completion as a fixed 
effects regression (including fixed effects for country and year).

```{r}
wdi_fixedeffects <- plm(GDP.avg ~ secondary.complet.avg, data=wdi_data, 
                 index=c("country","year"), model="within", effect="twoways")

summary(wdi_fixedeffects)

wdi_fixedeffects2 <- plm(GDP.avg ~ secondary.complet.avg + year, data=wdi_data, 
                 index=c("country","year"), model="within")
summary(wdi_fixedeffects2)


```

### e. Autocorrelation robust inference

Test the coefficients of the model using the Stata-style $HC_1$ estimation of  
Newey-West heteroskedasticity and autocorrelated (HAC) robust standard errors.

- Use the function [vcovNW()](https://www.rdocumentation.org/packages/plm/versions/1.6-5/topics/vcovNW) 
for specifying the variance method inside of [coeftest()](https://www.rdocumentation.org/packages/lmtest/versions/0.9-35/topics/coeftest).

```{r}
coeftest(wdi_fixedeffects,vcov = vcovNW(wdi_fixedeffects,type="HC1"))

```


