---
title: "Introduction to Data Analysis in R"
subtitle: "Module 3: Programming, joining data, and more"
author: "Andrew Proctor" 
institute: "andrew.proctor@phdstudent.hhs.se"
date: "January 28, 2019"
output: 
  beamer_presentation:
      theme: "Amsterdam"
      highlight: espresso
      slide_level: 2
      toc: true
header-includes:
  - \usepackage{color}  
  - \usepackage{hyperref}
  - \definecolor{cornflower}{HTML}{1c86ee}
  - \hypersetup{
    colorlinks=true,
    linkcolor=cornflower,
    filecolor=magenta,      
    urlcolor=blue,
    }

---

```{r setup, include=FALSE}
library(tidyverse)
library(rio)
library(magrittr)
library(rvest)
library(glue)
```

# Intro

## Goals for Module

1. Basics of programming in R---learn how to write and use:
    + Iterations (loops and map functions)
    + Conditional statements
    + Basic functions
2. Learn how to perform different types of dataset joins.
3. Learn how to manipulate strings and use "regular expressions""
4. Learn basic web scraping



# **Revisiting basics**

## Assignment Operator

So far, when changing a data object, we have always been a bit repetitive:
```{r,eval=FALSE}
mydataframe <-  mydataframe %>% 
  rename(NewVarName = OldVarName)
```


Along with the standard pipe **(%>%)**, by loading the 
[magrittr](https://magrittr.tidyverse.org/) package, you can also use the 
so-called "assignment pipe" **(%<>%)**.

  - The above rename with the assignment pipe appears as:
  
```{r,eval=FALSE}
mydataframe %<>% rename(NewVarName = OldVarName)
```



## Lists

```{r, include=FALSE}
library(Ecdat)
library(tidyverse)
data(Males)
set.seed(28133)
wages_df <- Wages1 %>% as.tibble() %>% sample_n(3) %>% as.data.frame() %>% select(wage,school, sex, exper) %>% rename(schooling = school) %>% mutate(wage=exp(wage))

colors <- c("red","green","blue")
description <- "Data on wages and date information."
date_df <- tibble(month = c("Jan","Feb","March"), 
                   year = c(2017,2018,2019))
date_df %<>% as.data.frame()
# Source: Verbeek, Marno (2004) A Guide to Modern Econometrics, John Wiley and Sons.
```

Another subtlety glossed over so far are *lists.* 

  - As mentioned in module 1, vectors come in two forms: atomic vectors (with a 
  single data type) and lists (with heterogenous data types).
  
  - Lists can take as inputs not only single-valued elements, but also vectors 
  or data frames.
  
  - Creating a list from other objects is done with the **list()** function.
  
## List Creation Example  

```{r}
wages_df; date_df; description
mylist <- list(wages_df,date_df,description)
```

Where *wages_df* and *date_df* are data frames and description is a single character
element.

## List Creation Example ctd

\includegraphics[scale=0.5]{mylist.png}

## Subsetting a list

- To subset a vector/matrix/data frame, one uses single brackets, eg mydf[,].

- To refer to an object of a list, use double brackets.  

```{r}
mylist[[3]]
```

**Note:** The function **list()** does not take transfer the names of the data 
frames, so you will need to either subset by position or assign names to the 
list objects.

## Extracting a list

An easy way of extracting an object from a list is with the **extract2()** function 
from \href{https://magrittr.tidyverse.org/}{magrittr}. This allows you to extract a 
given list object by name or position.

```{r}
wage_data <- mylist %>% extract2(1)
wage_data
```


## The unlist function

Instead of creating more complicated data objects, **unlist()** takes a list
and turns it into a simple (atomic) vector.

**Example:**
```{r, include=FALSE}
simple_list <-list(1,2,3,4)
```

```{r}
str(simple_list)
simple_list %<>% unlist() %>% str()
```


# **Iteration**

## For loops

For tasks that you want to iterate over multiple data frames/variables/elements, you 
may want to think about creating a **loop**.

- A loop performs a function/functions multiple times, across either a list of objects
or a set of index values.

**Syntax:**
```{r, eval=FALSE}
for(indexname in range) {
  do stuff
}
```

## For loop across numeric values

```{r loopexample}
for (i in 1:4){
  print(i^2)
}

```


## For loop across named elements

You can also loop over elements instead of values.

- In the last module exercises, you had to convert the type of many variables. 
Here's one way you could do that with a loop:

```{r asnumeric}
nlsy97 <- import("nlsy97.rds")
factor.vars <- c("personid","year","sex","race",
                 "region","schooltype")
  for (i in factor.vars){
   nlsy97[,i] %<>% unlist() %>% as.factor()
  }

```

## The map() function

For iterations over vectors and dataframes, the 
\href{https://www.rdocumentation.org/packages/purrr/versions/0.2.4/topics/map}{map()}
function is a great alternative to the for loop.

Map functions take a user-supplied function and iterate it over:

- Elements for a vector

- Objects of a list

- Columns of a data frame

Map functions are much simpler to write than loops and are also generally a good 
bit faster.

- ***Sidenote***: Map is a part of the tidyverse collection of packages.  In base R, the \href{https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/apply}{apply()}
family of functions does roughly the same thing, but \href{https://www.rdocumentation.org/packages/purrr/versions/0.2.4/topics/map}{map()}
simplifies and improves this task.

## Using the map() function

**Syntax:**

```{r, eval=FALSE}
map(data, fxn, option1, option2...)
```

**Example:**

```{r}
nlsy97[,factor.vars] %<>% map(as.numeric) 
```

## Using class-specific map variants

There are multiple map variants that enforce a given data type on results.  You should 
use these whenever you want output of a certain class.

- **map_lgl** for logical vector

- **map_dbl** for numeric vector

- **map_chr** for character vector

- **map_df** for a data frame

## Example of difference with class-specific map variants

```{r}
nlsy.sub <- nlsy97 %>% select(parentincome, 
                              motheredyrs, gpa)
nlsy.sub %>% map_dbl(IQR, na.rm=TRUE)
nlsy.sub %>% map(IQR, na.rm=TRUE)
```

## Using map() with anonymous functions
\href{https://www.rdocumentation.org/packages/purrr/versions/0.2.4/topics/map}{map()} 
works with not only predefined functions, but also "anonymous functions"--- 
unnamed functions defined inside of map().

- Suppose I want the z-standardized values of the variables from the previous example:

```{r}
# Create Z Transform
ztransform <- map_df(nlsy.sub, function(x) 
  (x - mean(x,  na.rm=TRUE)) / sd(x, na.rm=TRUE)
  )
```

## Using map() with anonymous functions ctd

Did my anonymous function work? 

```{r}
# Means
map_dbl(ztransform, function(x) 
  round(mean(x, na.rm=TRUE),10))
#  Standard deviations
map_dbl(ztransform, function(x) 
  round(sd(x, na.rm=TRUE),10))
```

# **Conditionals**

## If statements

"If statements" are also a useful part of programming, either in conjunction 
with iteration or seperately.

- An if statement performs operations only if a specified condition is met.
    + An important thing to know, however, is that if statements evaluate conditions of
    length one (ie non-vector arguments).
    + We will cover a vector equivalent to the if statement shortly.

**Syntax**

```{r, eval=FALSE}
if(condition){
     do stuff
    }
```

## Example of an if statement

- In the for loop example, the loop was indexed over only the columns of indicator 
codes.  

- Equally, the loop could be done over all columns with an if-statement to 
change only the indicator codes.

```{r}
for (j in colnames(nlsy97)){
  
  if(j %in% factor.vars){
      nlsy97[,j] %<>% unlist() %>% as.numeric()
  }
}

```

## Multiple conditions

You can encompass several conditions using the  <span style="color:blue">**else if**</span> 
and catch-all <span style="color:blue">**else**</span> control statements.

```{r, eval=FALSE}
if (condition1) {
do stuff
} else if (condition2) {
do other stuff
} else {
do other other stuff
}
```

## Vectorized if statements

- As alluded to earlier, if statements can't test-and-do for vectors, but only 
single-valued objects.

- Most of the time, you probably want to use conditional statements on vectors. 
The vector equivalent to the if statement is \href{https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/ifelse}{ifelse()}

**Syntax:**
```{r, eval=FALSE}
ifelse(condition, true_statement, false_statement)
```

The statements returned can be simple values, but they can also be functions or even
further conditions.  You can easily nest multiple \href{https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/ifelse}{ifelses} 
if desired.

## An ifelse example

```{r}
 numbers <- sample(1:30, 7); numbers
 ifelse(numbers %% 2 == 0,"even","odd")
```

**Note:** What if we tried a normal if statement instead?
```{r, warning=FALSE}
 if(numbers %% 2 == 0){
   print("even")} else{
     print("odd")}
```

## Multiple vectorized if statements

A better alternative to multiple nested `ifelse` statements is the tidyverse
**[case_when](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/case_when)**
function.

**Syntax:**

```{r, eval=FALSE}
case_when(
  condition1 ~ statement1,
  condition2 ~ statement2,
  condition3 ~ statement3,
)
```

## A case_when example

```{r}
nums_df <- numbers %>% as.tibble() %>% 
  mutate(interval = case_when(
  (numbers > 0 & numbers <= 10) ~ "1-10",
  (numbers > 10 & numbers <= 20) ~ "10-20",
  (numbers > 20 & numbers <= 30) ~ "20-30"))
nums_df[1:4,]
```

# **Functions**

## When you should write a function

If you find yourself performing the same specific steps more than a couple of 
times (perhaps with slight variations), then you should consider writing a function.

A function can serve essentially as a wrapper for a series of steps, where you 
define generalized inputs/arguments.

## Writing a function

**Ingredients:**

- Function name

- Arguments

- Function body

**Syntax:**

```{r, eval=FALSE}
function_name <- function(arg1, arg2, ...){
  do stuff
}
```

## Function example

Let's turn the calculation of even or odd that was completed earlier into a 
function:
```{r}
# Make odd function
odd <- function(obj){
   ifelse(obj %% 2 == 0,"even","odd")
}
```

***Notice*** that *obj* here is a descriptive placeholder name for the data object 
to be supplied as an argument for the function.

```{r}
odd(numbers)
```

## RStudio's "Extract Function"

A useful way of writing simple functions when you've already written the code 
for a specific instance is to use RStudio's *Extract Function* 
option, which is available from the code menu.

- *Extract function* will take the code chunk and treat any data objects referenced 
but not created within the chunk as function arguments.

# **Joins**

## Merging data

*Shifting gears from programming...*

Another staple task in applied work is combining data from multiple data sets. 
The tidyverse set of packages includes several useful types of merges (or "joins"):

- **[left_join()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join)** 
Appends columns from dataset B to dataset A, keeping all 
observations in dataset A.

- **[inner_join()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join)** 
Appends columns together, keeping only observations that appear 
in both dataset A and B.

- **[semi_join()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join)** 
Keeps only columns of dataset A for observations that appear 
in both dataset A and B.

- **[anti_join()](https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join)** 
Keeps only columns of dataset A for \ \ \ observations that *do not* 
appear in both dataset A and B.

## Joining using keys

The starting point for any merge is to enumerate the column or columns that 
uniquely identify observations in the dataset.

- For cross-sectional data, this might be a personal identifier or 
(for aggregate data) something like municipality, state, country, etc.

- For panel data, this will typically be both the personal/group identifier and 
a timing variable, for example Sweden in 2015 in a cross-country analysis.

## Mismatched key names across datasets

Sometimes the names of the key variables are different across datasets.

- You could of course rename the key variables to be consistent.

- But mismatched key names are easily handled by the tidyverse join functions.

**Syntax:**

```{r, eval=FALSE}
join_function(x, y, by = c("x_name" = "y_name"))
```

## left_join

The \href{https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join}{left\_join()} 
is the most frequent type of join, corresponding to a standard **merge** in Stata. 

- left_join simply appends additional variables from a second dataset to a main 
dataset, keeping all the observations (rows) of the first dataset.

**Syntax:**
```{r,eval=FALSE}
left_join(x, y, by = "key")
```

If the key is muliple columns, use <span style="color:blue">**c()**</span> to list them.

## left_join example
```{r,include=FALSE}
person_id <-c("001","002","003")
wage <- c(150,90,270)
earnings <- data.frame(person_id,wage)
schooling <-c(12,8,16)
person_id <-c("001","003","004")
educ <- data.frame(person_id,schooling)
```

```{r, warning=FALSE}
# Look at the datasets
earnings
educ
```

## left_join example ctd
```{r}
# Combine data
combined_data <- left_join(earnings, educ, 
                           by="person_id")

# Print data
combined_data
```

***Notice*** that schooling is equal to NA for person '002' because \ \ \ that person does 
not appear in the *educ* dataset.

## inner_join

If you want to combine the variables of two data sets, but only keep the observations 
present in both datasets, use the 
\href{https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join}{inner\_join()} 
function.

```{r, warning=FALSE}
combined_data <- inner_join(earnings, educ, 
                            by="person_id")
combined_data
```

## semi_join

To keep using only the variables in the first dataset, but where observations in 
the first dataset are matched in the second dataset, use 
\href{https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join}{semi\_join()}.

- semi_join is an example of a *filtering join*. 
Filtering joins don't add new columns, but instead just filter observations 
for matches in a second dataset.
- left_join and inner_join are instead known as *mutating joins*, because new 
variables are added to the dataset.

## semi_join example

```{r, warning=FALSE}
filtered_data <- semi_join(earnings, educ, 
                           by="person_id")
filtered_data
```

## anti_join

Another *filtering join* is 
\href{https://www.rdocumentation.org/packages/dplyr/versions/0.7.3/topics/join}{anti\_join()}, 
which filters for observations that are *not matched* in a second dataset.

```{r, warning=FALSE}
filtered_data <- anti_join(earnings, educ, 
                           by="person_id")
filtered_data
```

There are still other join types, which you can read about \href{http://dplyr.tidyverse.org/reference/join.html}{here}.

## Appending data

Finally, instead of joining different datasets for the same individuals, sometimes 
you want to join together files that are for different individuals within the same
dataset.

- When join data where the variables for each dataset are the same, but the observations 
are different, this is called *appending* data.

The function for appending data in the tidyverse is:
```{r, eval=FALSE}
bind_rows(list(dataframe1,dataframe2,...))
```

# **Manipulating text**

## Concatenating strings

The last type of data preparation that we will cover in this course is manipulating 
string data.
  
  - The simplest string manipulation may be concatenating (ie combining) strings.
  
      + A great function for combining string in R is the **glue()** function, part 
      of the Tiydverse \href{https://glue.tidyverse.org/reference/glue.html}{glue package}.
      
  - The glue function lets you reference variable values inside of text strings 
  by writing the variable in curly brackets **{}** inside of the string.
  
## Glue Example

```{r}
date_df %<>% mutate(
   say.month = glue("The month is {month}"),
   mo.yr = glue("{month} {year}")
 )
date_df
```

## Glue Example 2

```{r}
numbers <- c(1,2,3)
for (i in numbers){
  print(glue("The magic number is {i}"))
}
```

## Extracting and replacing parts of a string

Other common string manipulating tasks include extracting or replacing parts 
of a string.

  - These tasks can be done via the **str_extract()** and 
  **str_replace()** in the Tidyverse \href{https://stringr.tidyverse.org/index.html}{stringr} 
  package.

  - We saw examples of these two functions in the last seminar exercise:

##  Extracting and replacing parts of a string

The syntax for each function is:
```{r, eval=FALSE}
str_extract(string_object, "pattern_to_match")
str_replace(string_object, "pattern_to_match",
            "replacement_text")

```

By default, both function operate on the first match of the specified pattern.  To 
operate on *all* matchs, add "_all" to the function name, as in:

```{r, eval=FALSE}
str_extract_all(string_object, "pattern_to_match")

```

## Extract and replace example

In the last seminar, we created a "year" column from years indicated in the "variable" 
column text via the expression:

```{r, eval=FALSE}
nlsy97$year <- str_extract(nlsy97$variable, "[0-9]+")
```

After creating the "year" column, we then removed the year values from the values 
of the "variable" column by replacing these numbers with an empty string.

```{r, eval=FALSE}
nlsy97$variable <- str_replace(nlsy97$variable, 
                               "[0-9]+","")
```

## Trimming a string

When working with formatted text, a third common task is to remove extra spaces 
before or after the string text.

- This is done with the **str_trim()** function.  The syntax is:

```{r, eval=FALSE}
str_trim(string, side = left/right/both)
```

Note, when printing a string, any formatting characters are shown.  To view how 
the string looks formatted, use the **ViewLines()** function.

## Using regular expressions with strings

Often we want to modify strings based on a pattern rather than an exact 
expression, as seen with the **str_extract()** and **str_replace()** examples.

  - Patterns are specified in R (as in many other languages) using a syntax 
  known as "regular expressions" or *regex*.
  
  - Today, we will very briefly introduce some regular expressions.
 
## Common Expressions

- To match "one of" several elements, refer to them in square brackets, eg: [abc]
- To match one of a range of values, use a hyphen to indicate the range: e.g. [a-Z],[0-9]
- To match either of a couple of patterns/expressions, use the OR operator, eg: "2017|2018"
- There are also abbreviation for one of specific types of characters
  
    + eg: **[:digit:]** for numbers, **[:alpha:]** for letters, **[:punct:]** for punctuation, and $\textbf{.}$ for every character.

    + See the RStudio cheat sheet on stringr for more examples \ \ \ \ \ (and in general,
    as a brilliant reference to *regex*)
 
## How many times to match?

Aside from specifiying the characters to match, such as "[0-9]", another important 
component of regular expressions is how many time should the characters appear.

- "[0-9]" will match any part of a string composed of exactly *1* number.
- "[0-9]+" will match any part of a string composed of *1 or more* numbers.
- "[0-9]{4}" will match any part of a string composed of exactly *4* numbers.
- "[0-9]*" will match any part of a string composed of zero or more numbers.

## Examples with repetition

```{r, eval=FALSE, include=FALSE}
str_extract(messy_var, "^(19|20)*[0-9]{2}$")
```
Suppose we want to extract year data that is mixed in with other data as well.
```{r}
messy_var <- c(1,1987,2006,2010,307,2018)
str_extract(messy_var, "[0-9]")
str_extract(messy_var, "[0-9]+")
str_extract(messy_var, "[0-9]{4}")
```

## Escaping special characters

Often, special characters can cause problems when working with strings. For 
example, trying to add a quote can result in R thinking you are trying to close 
the string.

For most characters, you can "escape" (cause R to read as part of the string) 
special characters by prepending them with a backslash.

**Example:**
```{r}
quote <- "\"Without data, you're just another person 
with an opinion.\" - W. Edwards Deming."
writeLines(quote)
```

## Matching strings that precede or follow specific patterns

To match part of a string that occurs before or after a specific other pattern,
you can also specify "lookarounds", the pattern the match should precede or follow:

To match a string pattern x, preceded or followed by y:
  
  - **y precedes x:** "(?<=y)x"
  
  - **y follows x:** "x(?=y)"

## Look around example

```{r}

price_info <-c("The price is 5 dollars")
str_extract(price_info, "(?<=(The price is )).+")
str_extract(price_info, ".+(?=( dollars))")

```

# **Web Scraping**

## Web scraping with Rvest
"Scraping" data from the web - that is, automating the retrieval of data displayed
online (other than through API) is an increasingly common data analysis task.

- Today, we will briefly explore very rudimentary web scraping, using the **rvest** 
package.

- The specific focus today is only on scraping data structued as a table on 
a webpage.  The basic method highlighted will work much of the time - but does not 
work for every table.

## Using rvest to scrape a table

- The starting point for scraping a web table with rvest is the **read_html()** 
function, where the URL to the page with data should go.

- After reading the webpage, the table should be parsed.  For many tables, 
the **read_html** can be piped directly into the **html_table()**  function.

    + If this works, the data should then be converted from a list into a 
    dataframe/tibble.

- If **html_table()** does not work, a more robust option is to first pipe **read_html** 
into **html_nodes(xpath = "//table")** and then into **html_table(fill=TRUE)**
  
    + **html_nodes(xpath = "//table")** looks for all HTML \ \ \ \ \ \ objects coded as a 
    table, hence tends to produces lists with several objects.


## Web scraping example

```{r, include=FALSE}
tech_stock_names <- c("MSFT","AMZN","GOOGL","AAPL",
                      "FB","INTC","CSCO")
tech_stocks <- list()
for(j in 1:length(tech_stock_names)){
tech_stocks[[j]] <-read_html(
glue(
"https://finance.yahoo.com/quote/{tech_stock_names[j]}/history")) %>%
html_table() %>% as.data.frame() %>% 
  mutate(stock = tech_stock_names[j])
}
```

```{r, eval=FALSE}
tech_stock_names <- c("MSFT","AMZN","GOOGL","AAPL",
                      "FB","INTC","CSCO")
tech_stocks <- list()
for(j in 1:length(tech_stock_names)){
  tech_stocks[[j]] <-read_html(
  glue("https://finance.yahoo.com/quote/{tech_stock_
       names[j]}/history")) %>%
  html_table() %>% as.data.frame() %>% 
    mutate(stock = tech_stock_names[j])
}
```

## Web scraping example ctd

\includegraphics[scale=0.5]{tech_stocks_list.png}

## Web scraping example

```{r}
tech_stocks %<>% bind_rows()
tech_stocks[1:5,c(1,6:8)]
```

## Another webscraping example

```{r}
gini_list <-read_html(
  "http://wdi.worldbank.org/table/1.3") %>%
  html_nodes(xpath ="//table") %>% 
  html_table(fill=TRUE) 
gini_data <- gini_list %>% extract2(3) %>% 
  as.data.frame() %>% select(1:3)
gini_data[1:3,]
```

 