---
title: "Module 3 Exercise"
output: 
  pdf_document:
    df_print: kable
    highlight: haddock
urlcolor: blue

---

```{r setup, include=FALSE}
library(rio)
library(tidyverse)
```


## Instructions

### Download and import the data

For this exercise, you are going to primarily be working with data from the US 
Bureau of Labor Statistics *Consumer Expenditure Survey* (CEX).  You can download 
the data [here](https://www.bls.gov/cex/pumd_data.htm#sas).

- Download the SAS **2016 Interview** file. Extract the data files and keep the 
following two files (moving them into a project folder):
    + *fmli161x.sas7bdat*, which corresponds to 2016Q1 data about household 
    characteristics and income.
    + *mtbi161x.sas7bdat*, which corresponds to 2016Q1 data about household 
    expenditures.

- Set the working directory.

- Import the two data files, calling them *characteristics* and *expenditures*.  You may 
want to convert them to tibbles.

```{r, include=FALSE}
characteristics <- import("./data/fmli161x.sas7bdat") %>% as.tibble()
expenditures <- import("./data/mtbi161x.sas7bdat")%>% as.tibble()
```

### Iterate a function over column names

- Change all the column names to lower case by iterating the function  [tolower()](https://www.rdocumentation.org/packages/h2o/versions/2.8.4.4/topics/tolower) 
over the column names of the two data frames.

```{r, include=FALSE}
colnames(characteristics) <- colnames(characteristics) %>% map(tolower)
colnames(expenditures) <- colnames(expenditures) %>% map(tolower)
```


### Keep specific columns

As you may've noticed, there's a lot (over 800) variables in the dataset.  Let's 
reduce this only to the following variables:

- **For characteristics:** \ \ \  newid, hh_cu_q,educ_ref, creditx,region, fincbtxm

- **For expenditures:** \ \ \  newid, cost, ref_mo, ref_yr

```{r, include=FALSE}
characteristics <- characteristics %>% select(newid,hh_cu_q,educ_ref,creditx, region, fincbtxm)
expenditures <- expenditures %>% select(newid,cost,ref_mo,ref_yr)
```

### Rename columns

Rename the following variables:

- *hh_cu_q* to *hh_size*

- *fincbtxm* to *hh_income*

```{r, include=FALSE}
characteristics <- characteristics %>% rename(hh_size=hh_cu_q)
characteristics <- characteristics %>% rename(hh_income=fincbtxm)

```

### Change the class of columns

Make all the variables except for *newid* into numeric for both data frames, 
using a loop or map function.

```{r, include=FALSE}
characteristics[,-1] <- characteristics[,-1] %>% map(as.numeric)
expenditures[,-1] <- expenditures[,-1] %>% map(as.numeric)

```


### Sample 80% of observations for both datasets

- To make the joins in the next step a little more interesting, first modify the datasets 
so that they are only a 80% sample of the full datasets.

```{r, include=FALSE}
characteristics <- sample_frac(characteristics, 0.8)
expenditures <- sample_frac(expenditures, 0.8)
```

### Aggregate expenditures by household

For both datasets, *newid* is a unique identifier for household.  In the *expenditures*
dataset, each expenditure is entered in seperately so that each household shows up 
many times.  Using the appropriate tidyverse functions, sum up the expenditures by 
 *newid*, replacing *expenditures* with this aggregated information. 

### Practice different joins

Try using the different *join* functions covered in the module. 

- In particular, first perform a traditional join that keeps all of the 
observations from *expenditures* and the columns of *expenditures* and 
*characteristics*. Save the result of this join as *cex_data*.

- Also try a join that keeps the columns of *expenditures* and *characteristics*, 
but only the observations in both datasets.

- Try a join that keeps only the columns of *expenditures*, with only the 
observations of *expenditures* that are not matched in *characteristics*.


```{r, include=FALSE}
cex_data <- left_join(expenditures, characteristics, by="newid")
```

### Use conditional statements to create indicators for region

Starting from *cex_data*, create indicators for each region value. You might find the 
[unique()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/unique) 
function helpful.

```{r, include=FALSE}
cex_data <- cex_data %>% mutate(region1 = ifelse(region == 1,1,0),
                    region2 = ifelse(region == 2,1,0),
                    region3 = ifelse(region == 3,1,0),
                    region4 = ifelse(region == 4,1,0),
                    region4 = ifelse(is.na(region),1,0)
                    )

```



### Write your own simple linear regression function

Finally, try your hand at writing a function.  In particular, try to write a 
function that produces the coefficient in a linear regression. In matrix notation, 
the formula for $\hat{\beta}_{OLS}$ is:
$$ \hat{\beta}_{OLS} = (X'X)^{-1}(X'y) $$
You will need some more matrix multiplication operators for this:

- [solve(A)](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/solve) 
    yields the inverse of matrix A.
    
- [t()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/t) 
    provides the transpose of matrix A. 

Also, remember to add a column of ones to include an intercept in the model. You 
can make a vector of ones by using the 
[rep()](https://www.rdocumentation.org/packages/base/versions/3.4.3/topics/rep) 
inside of vector or matrix definition.

Once you've finished writing the function, try running it to produce the parameter 
estimates from a regression of expenditures on any of the other variables in *cex_data*.
